Something about Continuous Deductive Reasoning over Streamed RDF Data, maybe claim Soundness or some other Theoretical Quality

### 2015-06-25 content/contribution/continuous-datalog.tex
added a few paragraphs and an axiom regarding instance entailment, allowing the proof to work without proving from the previous axioms that E_i \cap E_{i-R} |= {}

	Continuous Datalog is the extension of positive Datalog with stream-processing semantics, with a view to accommodating the aims of the latter within the well-defined semantics of the former, which are shared with RIF-Core and therefore support reasoning at the expressivity of OWL 2 RL.
	This comprises the extension of Datalog programs to include sources of continuously generated/reported facts, and the application of sliding window semantics over those sources, thereby providing for the interpretation of Continuous Datalog programs as sequences of instantaneous states that are each valid positive Datalog programs.
	As such, these states each inherit the semantics of positive Datalog.
	Furthermore, it involves the continuous derivation of entailments from this sequence of instantaneous states as a set of streams of facts;
	these streams share a data model with the streams of continuous queries, as defined by the stream processing community, thereby allowing the integration of Continuous Datalog programs into stream processing pipelines.
	In addition, these streams may be trivially manipulated to retrieve the instantaneous sets of entailments produced by the more semantically well-defined offerings from the reasoning community, as proven in \refsec{semantics: proof for entailments}.

@@

@@

			Such a continuous program entails a triple composed of the entailments of ${A \vDash A'}$, a stream of entailments $P$ and a stream of entailment negations $N$, whose contents are defined in Axioms~\ref{axiom:continuous datalog: positive window increment},~\ref{axiom:continuous datalog: negative window increment} and~\ref{axiom:continuous datalog: entailment precedes negation}.

@@

			The truth value of a ${CDP}$ is instantaneous, with the the truth function of Continuous Datalog $T^{CD}$ taking both a program ${CDP}$ and an instant ${t \in \mathbb{N}}$ as arguments.

@@

			The state of a ${CDP}$ at a given instant ${t \in \mathbb{N}}$ (${CDP_{i}}$) is the union of the axioms $A$ and the window $W^{S,R}_{i}$.
			As both $A$ and $W^{S,R}_{i}$ are sets of Datalog axioms, a ${CDP_{i}}$ constitutes a valid Datalog program that may be evaluated by the truth function of Datalog $T^{D}$, the result of which is the definition for the instantaneous truth of the ${CDP}$ at instant $t$ by $T^{CD}$.

@@

				T^{CD} \left( CDP, t \right) = T^{D} \left( CDP_{i} \right)

@@

			The window of range 1 at any given instant ${t \in \mathbb{N}}$ over the stream $P$ entailed by a ${CDP}$ is defined as the set of facts that are entailed by the state of the ${CDP}$ at instant $t$ that were not entailed by the state of ${CDP}$ at the previous instant instance ${t-1}$.

@@

			The window of range 1 at any given instant ${t \in \mathbb{N}}$ over the stream $N$ entailed by a ${CDP}$ is defined as the set of facts that were entailed by the state of the ${CDP}$ at the previous instant instance ${t-1}$ that are no longer entailed by the state of ${CDP}$ at instant $t$.

@@

			As such, a window over the stream of entailments $P$ at any given instant ${t \in \mathbb{N}}$ will be disjoint from the 

@@

				\forall Y \geq 1 \left( W^{P,1}_{i} \cap W^{N,1}_{i-Y} = \varnothing \right)

@@

	From this, we can see that each instance of an entailment in some $E_{i}$ may take the same form as an instance of any other fact in a stream of facts, being ${p(s,o):t}$, where $t$ is the real time associated with the instant $i_{e}$ in which ${p(s,o)}$ becomes entailed by the ${CDP}$
	(i.e. when it becomes newly justified by some subset of instances in $W^{S,R}_{i_{e}}$).
	As such, the stream of entailments $P$ of a ${CDP}$ takes the same form as the \emph{streams of facts} already described, thus producing a concise stream of entailments that may be read by further stream processing.
	In addition, the stream of negations $N$ of a ${CDP}$ also takes this form, but differs from the streams discussed thus far in that it is not ordered by the entailment time $i_{e}$ of the entailment instances but by the time at which their justification \emph{ceases} to be valid
	(i.e. the first instant ${i_{n} \geq i_{e}}$ in which $W^{S,R}_{i_{n}}$ no longer contains the subset of fact instances first present in $W^{S,R}_{i_{e}}$ that justified the entailment instance ${p(s,o):t}$).

	Thus far the translation from instantaneous sets of entailments to streams of entailments has been stated axiomatically as the definition of those streams.
	However, it is proven in \ref{sec:semantics: proof for entailments}, from these axioms, that it is also possible to reconstruct exactly the instantaneous set $E_{i}$ of entailment instances of any ${CDP}$ at any given instant from the windows of sufficient range over its entailed streams $P$ and $N$ at that instant.
	\[ \forall Y \geq X \geq R \, \left( E_{i} = W^{P,X}_{i} \setminus W^{N,Y}_{i} \right) \]
	The minimum range $X$ is such to achieve the \emph{complete} set of entailment instances $E_{i}$ of the ${CDP}_{i}$.
	In addition, the minimum range $Y$ is such to achieve the \emph{correct} set $E_{i}$.
	As such, a \emph{valid window} that contains only and exactly those instances of entailments in $E_{i}$ at every instant $i$ may be trivially maintained from the streams $P$ and $N$ continuously entailed by some ${CDP}$.
\end{nestedsection}

>>>>

	Continuous Datalog is the extension of positive Datalog with stream-processing semantics, with a view to accommodating the aims of the latter within the well-defined semantics of the former, which are shared with RIF-Core and therefore support reasoning at the expressivity of OWL 2 RL.
	This comprises the extension of Datalog programs to include sources of continuously generated/reported facts, and the application of sliding window semantics over those sources, thereby providing for the interpretation of Continuous Datalog programs as sequences of instantaneous states that are each valid positive Datalog programs.
	As such, these states each inherit the semantics of positive Datalog.
	Furthermore, it involves the continuous derivation of entailments from this sequence of instantaneous states as a set of streams of facts;
	these streams share a data model with the streams of continuous queries, as defined by the stream processing community, thereby allowing the integration of Continuous Datalog programs into stream processing pipelines.
	In addition, in can be shown that these streams may be trivially manipulated to retrieve the instantaneous sets of entailments produced by the more semantically well-defined offerings from the reasoning community.

@@

	Traditionally, a set of axioms $A$ entails another set of axioms $E$ in which the existence of one fact in $E$ may be justified by many distinct subsets of $A$.
	When considering streams of facts, in which one fact may occur multiple times and be considered distinct based on its entailment time, temporal information would be lost if sets of axioms including fact instances (denoted $A_{inst}$) were to entail sets of axioms without.
	As such, where a fact would be entailed by some set of facts $A$ that are instantiated on some stream, an instance of that fact should be entailed by the set of instances of those facts present on that stream.
	Furthermore, each such justification in $A_{inst}$ for a fact entails a unique instance of that fact, such that each entailed instance has a single justification within the original set of axioms, and many instances of the entailed fact may occur in the entailed set of axioms $E_{inst}$.
	This form of entailment shall be henceforth referred to as ``instance entailment'', and shall be the norm when discussing the entailments of sets of axioms that include instances of facts.

	It should be noted that removing all temporal information from a set of axioms including fact instances $A_{inst}$ produces the set of traditional Datalog axioms $A$ instantiated in $A_{inst}$.
	This is be represented as the function ${trad\left( A_{inst} \right) = A}$.
	In addition, applying the ${trad()}$ function to both the original axiom instances $A_{inst}$ and the entailed axiom instances $E_{inst}$ maintains the relationship of entailment between the two sets:
	\[ \left( A_{inst} \vDash E_{inst} \right) \rightarrow \left( A \vDash E \right) \]

	From this description of \emph{instance entailment}, Axiom~\ref{axiom:continuous datalog: entailment disjointness} may be derived.
	\begin{enumerate}\setcounter{enumi}{\thecontinuousDatalogAxioms}
		\item\label{axiom:continuous datalog: entailment disjointness}
			If two sets, $A_{inst}$ and $B_{inst}$, of axiom instances are disjoint, then the any set $E_{inst}$ of axiom instances each justified at least in part by $A_{inst}$ is necessarily disjoint from any set $F_{inst}$ of axiom instances each justified at least in part by $B_{inst}$, because entailed fact instances are distinct with regards to the fact instances that justify them.
			\begin{multline*}
				\left( A_{inst} \cap B_{inst} = \varnothing \right) \wedge \left( A_{inst} \vDash E_{inst} \right) \\
					\wedge \left( B_{inst} \vDash F_{inst} \right) \rightarrow \left( E_{inst} \cap F_{inst} = \varnothing \right)
			\end{multline*}
		\setcounter{continuousDatalogAxioms}{\theenumi}
	\end{enumerate}

@@

			Such a continuous program entails a triple composed of the traditional entailments of ${A \vDash A'}$, a stream of entailed instances $P$ and a stream of instance negations $N$, whose contents are defined in Axioms~\ref{axiom:continuous datalog: positive window increment},~\ref{axiom:continuous datalog: negative window increment} and~\ref{axiom:continuous datalog: entailment precedes negation}.

@@

			The truth value of a ${CDP}$ is instantaneous, with the the truth function of Continuous Datalog $T^{CD}$ taking both a program ${CDP}$ and an instant ${i \in \mathbb{N}}$ as arguments.

@@

			The state of a ${CDP}$ at a given instant ${i \in \mathbb{N}}$ (${CDP_{i}}$) is the union of the axioms $A$ and the window $W^{S,R}_{i}$.
			As both $A$ and ${trad\left( W^{S,R}_{i} \right)}$ are sets of Datalog axioms, a ${CDP_{i}}$ constitutes a valid Datalog program that may be evaluated by the truth function of Datalog $T^{D}$, the result of which is the definition for the instantaneous truth of the ${CDP}$ at instant $i$ by $T^{CD}$.

@@

				T^{CD} \left( CDP, i \right) = T^{D} \left( CDP_{i} \right)

@@

			The window of range 1 at any given instant ${i \in \mathbb{N}}$ over the stream $P$ entailed by a ${CDP}$ is defined as the set of facts that are entailed by the state of the ${CDP}$ at instant $i$ that were not entailed by the state of ${CDP}$ at the previous instant instance ${i-1}$.

@@

			The window of range 1 at any given instant ${i \in \mathbb{N}}$ over the stream $N$ entailed by a ${CDP}$ is defined as the set of facts that were entailed by the state of the ${CDP}$ at the previous instant instance ${i-1}$ that are no longer entailed by the state of ${CDP}$ at instant $i$.

@@

			As such, a window of any range ${X \in \mathbb{N}}$ over the stream of entailments $P$ at any given instant ${i \in \mathbb{N}}$ is disjoint from a window of any range ${Y \in \mathbb{N}}$ over the the stream of entailment negations $N$ at any instant after ${i - X}$

@@

				\forall X' \geq X \left( W^{P,X}_{i} \cap W^{N,Y}_{i-X'} = \varnothing \right)

@@

	From this, we can see that each instance of an entailment in some $E_{i}$ may take the same form as an instance of any other fact in a stream of facts, being ${p(s,o):t}$, where $t$ is the real time associated with the instant $i_{e}$ in which ${p(s,o)}$ becomes newly justified by the ${CDP}$
	(more specifically, by some subset of the instances in $W^{S,R}_{i_{e}}$).
	As such, the stream of entailment instances $P$ of a ${CDP}$ takes the same form as the \emph{streams of facts} already described, thus producing a concise stream of entailments that may be read by further stream processing.
	In addition, the stream of negations $N$ of a ${CDP}$ also takes this form, but differs from the streams discussed thus far in that it is not ordered by the entailment time $i_{e}$ of the entailment instances but by the time at which their justification \emph{ceases} to be valid
	(i.e. the first instant ${i_{n} \geq i_{e}}$ in which $W^{S,R}_{i_{n}}$ no longer contains the subset of fact instances first present in $W^{S,R}_{i_{e}}$ that justified the entailment instance ${p(s,o):t}$).

	Thus far the translation from instantaneous sets of entailments to streams of entailments has been stated axiomatically as the definition of those streams.
	However, it can be shown from these axioms,
	as in \ref{sec:semantics: proof for entailments},
	that it is also possible to reconstruct exactly the instantaneous set $E_{i}$ of entailment instances of any ${CDP}$ at any given instant from the windows of sufficient range over its entailed streams $P$ and $N$ at that instant.
	\[ \forall Y \geq X \geq R \, \left( E_{i} = W^{P,X}_{i} \setminus W^{N,Y}_{i} \right) \]
	The minimum range $X$ is such to achieve the \emph{complete} set of entailment instances $E_{i}$ of the ${CDP}_{i}$.
	In addition, the minimum range $Y$ is such to achieve the \emph{correct} set $E_{i}$.
	As such, a \emph{valid window} that contains only and exactly those instances of entailments in $E_{i}$ at every instant $i$ may be trivially maintained from the streams $P$ and $N$ continuously entailed by some ${CDP}$.
\end{nestedsection}

### 2015-06-25 content/contribution/continuous-datalog/proof-of-entailment-semantics.tex
added a few paragraphs and an axiom regarding instance entailment, allowing the proof to work without proving from the previous axioms that E_i \cap E_{i-R} |= {}

\begin{nestedsection}{Entailment Semantics: Proof By Inductive Lemma}{semantics: proof for entailments}

	\begin{enumerate}\setcounter{enumi}{\thecontinuousDatalogAxioms}
		\item\label{axiom:continuous datalog: disjoint entailments}
			The instantaneous set of stream-justified entailments $E_{i}$ of a ${CDP = \{A,S,R\}}$ is necessarily disjoint with the instantaneous set of stream-justified entailments $E_{i-X}$ of the same ${CDP}$ when ${X \geq R}$.
			\begin{equation*}
				E_{i} \cap E_{i-X} = \varnothing
			\end{equation*}
		\setcounter{continuousDatalogAxioms}{\theenumi}
	\end{enumerate}

	\begin{hyp}\labelhyp{continuous datalog: entailment} \hfill \\
		For any ${CDP = \{A,S,R\} \vDash \{A',P,N\}}$, where ${CDP_{i} \vDash A' \cup E_{i}}$ for any ${i \in \mathbb{N}}$, the set of facts in a window of any range ${X \geq R}$ over the stream $P$ at a given instant ${i}$ and not in a window of any range ${Y \geq X}$ over the stream $N$ at the same instant $i$ is equal to the set of facts $E_{i}$ transitively entailed by the ${CDP}$ at that instant $i$:
		\begin{equation*}
			W^{P,X}_{i} \setminus W^{N,Y}_{i} = E_{i}
		\end{equation*}
	\end{hyp}
	In order to prove \refhyp{continuous datalog: entailment}, in which the windows over the entailed streams have a Range greater than or equal to that over the input streams, I present Lemma~\ref{hyp:continuous datalog: entailment lemma} as the general case where the windows have an arbitrary range.
	\begin{lem}\label{hyp:continuous datalog: entailment lemma} \hfill \\
		For any ${CDP \vDash \{A',P,N\}}$, where ${CDP_{i} \vDash A' \cup E_{i}}$ for any ${i \in \mathbb{N}}$, the set of facts in a window of any range ${X \in \mathbb{N}}$ over the stream $P$ at a given instant ${i}$ and not in a window of any range ${Y \geq X}$ over the stream $N$ at the same instant $i$ is equal to the set of transient entailments at that instant $i$ less the intersection of all previous sets of transient entailments from instant ${i-1}$ up to and including ${i-X}$:
		\begin{equation*}
			W^{P,X}_{i} \setminus W^{N,Y}_{i} = E_{i} \setminus \mathop{\cap}^{X}_{d=1} E_{i-d}
		\end{equation*}
	\end{lem}

	\begin{proof}[Proof of \refhyp{continuous datalog: entailment} using Lemma~\ref{hyp:continuous datalog: entailment lemma}]\hfill\\
		Assuming that Lemma~\ref{hyp:continuous datalog: entailment lemma} holds for every widow size ${X \geq R}$ for some ${CDP = \{A,S,R\}}$ and ${Y \geq X}$ at every instant ${i \in \mathbb{N}}$:
		\begin{align*}
			E_{i} & = W^{P,X}_{i} \setminus W^{N,Y}_{i} \\
			\text{Lemma~\ref{hyp:continuous datalog: entailment lemma}} & = E_{i} \setminus \mathop{\cap}^{X}_{d=1} E_{i-d} \\
			& = \mathop{\cup}^{X}_{d=1} \left( E_{i} \setminus E_{i-d} \right) \\
			& = \left( E_{i} \setminus E_{i-X} \right) \cup \mathop{\cup}^{X-1}_{d=1} \left( E_{i} \setminus E_{i-d} \right) \\
			\text{Axiom~\ref{axiom:continuous datalog: disjoint entailments}} & = E_{i} \cup \mathop{\cup}^{X-1}_{d=1} \left( E_{i} \setminus E_{i-d} \right) \\
			& = E_{i} \cup \left( E_{i} \setminus \mathop{\cap}^{X-1}_{d=1} E_{i-d} \right) \\
			& = E_{i} && \qedhere
		\end{align*}
	\end{proof}

>>>>

\begin{nestedsection}{Entailment Semantics: Proof By Inductive Lemma}{semantics: proof for entailments}
	\begin{hyp}\labelhyp{continuous datalog: entailment} \hfill \\
		For any ${CDP = \{A,S,R\} \vDash \{A',P,N\}}$, where ${CDP_{i} \vDash A' \cup E_{i}}$ for any ${i \in \mathbb{N}}$, the set of facts in a window of any range ${X \geq R}$ over the stream $P$ at a given instant ${i}$ and not in a window of any range ${Y \geq X}$ over the stream $N$ at the same instant $i$ is equal to the set of facts $E_{i}$ transitively entailed by the ${CDP}$ at that instant $i$:
		\begin{equation*}
			W^{P,X}_{i} \setminus W^{N,Y}_{i} = E_{i}
		\end{equation*}
	\end{hyp}
	In order to prove \refhyp{continuous datalog: entailment}, in which the windows over the entailed streams have a Range greater than or equal to that over the input streams, I present Lemma~\ref{hyp:continuous datalog: entailment lemma} as the general case where the windows have an arbitrary range.
	\begin{lem}\label{hyp:continuous datalog: entailment lemma} \hfill \\
		For any ${CDP \vDash \{A',P,N\}}$, where ${CDP_{i} \vDash A' \cup E_{i}}$ for any ${i \in \mathbb{N}}$, the set of facts in a window of any range ${X \in \mathbb{N}}$ over the stream $P$ at a given instant ${i}$ and not in a window of any range ${Y \geq X}$ over the stream $N$ at the same instant $i$ is equal to the set of transient entailments at that instant $i$ less the intersection of all previous sets of transient entailments from instant ${i-1}$ up to and including ${i-X}$:
		\begin{equation*}
			W^{P,X}_{i} \setminus W^{N,Y}_{i} = E_{i} \setminus \mathop{\cap}^{X}_{d=1} E_{i-d}
		\end{equation*}
	\end{lem}

	\begin{proof}[Proof of \refhyp{continuous datalog: entailment} using Lemma~\ref{hyp:continuous datalog: entailment lemma}]\hfill\\
		Assume that Lemma~\ref{hyp:continuous datalog: entailment lemma} holds for every widow size ${X \geq R}$ for some ${CDP = \{A,S,R\}}$ and ${Y \geq X}$ at every instant ${i \in \mathbb{N}}$.
		Remember that ${W^{S,R}_{i} \cap W^{S,R}_{i-R} = \varnothing}$, ${W^{S,R}_{i} \vDash E_{i}}$ and ${W^{S,R}_{i-R} \vDash E_{i-R}}$.
		Therefore:
		\begin{align*}
			E_{i} & = W^{P,X}_{i} \setminus W^{N,Y}_{i} \\
			\text{Lemma~\ref{hyp:continuous datalog: entailment lemma}} & = E_{i} \setminus \mathop{\cap}^{X}_{d=1} E_{i-d} \\
			& = \mathop{\cup}^{X}_{d=1} \left( E_{i} \setminus E_{i-d} \right) \\
			& = \left( E_{i} \setminus E_{i-X} \right) \cup \mathop{\cup}^{X-1}_{d=1} \left( E_{i} \setminus E_{i-d} \right) \\
			\text{Axiom~\ref{axiom:continuous datalog: entailment disjointness}} & = E_{i} \cup \mathop{\cup}^{X-1}_{d=1} \left( E_{i} \setminus E_{i-d} \right) \\
			& = E_{i} \cup \left( E_{i} \setminus \mathop{\cap}^{X-1}_{d=1} E_{i-d} \right) \\
			& = E_{i} && \qedhere
		\end{align*}
	\end{proof}
